Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-05-21 performance_comparison.tex
[MM] definitive constraints acronym

The first parameter analyzed is the factor $\ga^\textrm{OTC}$, which multiplies the standard deviation of the constraint, $\sigma^\textrm{OTC}$, to determine the total back-off term $\be^\textrm{OTC}$. This parameter directly influences the probability of satisfying the stay-on-track constraint: higher values of $\ga^\textrm{OTC}$ correspond to more conservative (i.e., robust) behavior. Specifically, $\ga^\textrm{OTC} = 0$ yields a satisfaction probability of 50\%, while $\ga^\textrm{OTC} = 3$ corresponds to 99\%. For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{OTC}$ within this range.

@@

\subsection{Robustified constraints comparison}
It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:adherence_constraint} and~\ref{sec:ontrack_constraint}---influences the driving style. This analysis compares the nominal trajectory, obtained without robustified constraints, with those resulting from a robustified stay-on-track constraint (denoted as OTC), a robustified adherence limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as OTC+FLC). All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{OTC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of meeting both constraints.

>>>>

The first parameter analyzed is the factor $\ga^\textrm{TLC}$, which multiplies the standard deviation of the constraint, $\sigma^\textrm{TLC}$, to determine the total back-off term $\be^\textrm{TLC}$. This parameter directly influences the probability of satisfying the stay-on-track constraint: higher values of $\ga^\textrm{TLC}$ correspond to more conservative (i.e., robust) behavior. Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\%, while $\ga^\textrm{TLC} = 3$ corresponds to 99\%. For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ within this range.

@@

\subsection{Robustified constraints comparison}
It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style. This analysis compares the nominal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC). All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of meeting both constraints.

### 2025-05-21 stochastic_vehicle_dynamics.tex
[MM] definitive constraints acronym

\subsection{Robust friction limit constraint formulation}
\label{sec:adherence_constraint}
The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces. As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tyre's adherence ellipse.

@@

\subsection{On-track constraint}
\label{sec:ontrack_constraint}
An on-track constraint is required to ensure that the vehicle remains within the circuit boundaries. The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:
\begin{equation}
	\begin{Bmatrix}
		x\\y
	\end{Bmatrix}
	-\boldsymbol{c} - e\bn=\boldsymbol{0} \label{eq:onplane_constraint}
\end{equation}
%\begin{equation}
%	\bp-\boldsymbol{c}_k - e_k\bn_k=\boldsymbol{0} \label{eq:onplane_constraint}
%\end{equation}
where $\boldsymbol{c}$ indicates the position of the centerline, and $\bn$ is the unit normal vector to the centerline. Equation~\eqref{eq:onplane_constraint} constraints the CoM of the vehicle to lie on the vertical plane defined by the unit normal vector $\bn$. The previously introduced algebraic variable $e$ is further constrained to take values between $e_\textrm{min}$ and $e_\textrm{max}$, which are determined based on the width of the circuit and the vehicle's front and rear tracks.

In order to take into account for the back-off terms, the maximum and minimum values are modified coherently, so that:
\begin{equation}
	e_\textrm{min} + \be^\textrm{OTC} \leq e \leq e_\textrm{max} - \be^\textrm{OTC}\label{eq:ontrack_constraint}
\end{equation}
where $\be^\textrm{OTC}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{OTC}$ depends solely on $\bn$.
%Adding a back-off to this constraint produces the same effect as an increase of the front and rear tracks of the vehicle does.

>>>>

\subsection{Robust friction limit constraint formulation}
\label{sec:FLC}
The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces. As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tyre's adherence ellipse.

@@

\subsection{Robust track limit constraint formulation}
\label{sec:TLC}
An on-track constraint is required to ensure that the vehicle remains within the circuit boundaries. The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:
\begin{equation}
	\begin{Bmatrix}
		x\\y
	\end{Bmatrix}
	-\boldsymbol{c} - e\bn=\boldsymbol{0} \label{eq:onplane_constraint}
\end{equation}
%\begin{equation}
%	\bp-\boldsymbol{c}_k - e_k\bn_k=\boldsymbol{0} \label{eq:onplane_constraint}
%\end{equation}
where $\boldsymbol{c}$ indicates the position of the centerline, and $\bn$ is the unit normal vector to the centerline. Equation~\eqref{eq:onplane_constraint} constraints the CoM of the vehicle to lie on the vertical plane defined by the unit normal vector $\bn$. The previously introduced algebraic variable $e$ is further constrained to take values between $e_\textrm{min}$ and $e_\textrm{max}$, which are determined based on the width of the circuit and the vehicle's front and rear tracks.

In order to take into account for the back-off terms, the maximum and minimum values are modified coherently, so that:
\begin{equation}
	e_\textrm{min} + \be^\textrm{TLC} \leq e \leq e_\textrm{max} - \be^\textrm{TLC}\label{eq:TLC}
\end{equation}
where $\be^\textrm{TLC}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{TLC}$ depends solely on $\bn$.
%Adding a back-off to this constraint produces the same effect as an increase of the front and rear tracks of the vehicle does.
